PLANETESIMAL ACCRETION AT SHORT ORBITAL PERIODS

### 2022-09-07 main.tex
Edits through section 5.4

We show the final state of the ``fdHi'' simulation in figure \ref{fig:fulldisk_e_m}. In the top panel,
the initial (contours) and final (points) state of the simulation is shown in the orbital period-eccentricity plane. The size of the points indicates the relative mass of the bodies. In the bottom panel, the mass of the largest bodies is shown as a function of orbital period. The solid curves indicate the isolation mass,
\begin{equation}\label{eq:iso}
	M_{iso} = \left[ \frac{\left( 2 \pi a^2 \Sigma \tilde{b} \right)^3}{3 M_{*}} \right]^{1/2},
\end{equation}
where $\tilde{b}$ is the size of the feeding zone in units
of Hill radii. For two bodies on circular, non-inclined orbits,  $\tilde{b} = 2
\sqrt{3}$ is the smallest orbital separation that produces a non-negative Jacobi energy
and permits a close encounter (see \citet{naka88}).
The isolation mass is plotted in blue and orange for these two values of $\tilde{b}$, respectively.

@@

In section \ref{sec:narrow}, we postulated that the increased importance of inelastic damping in the inner, non-oligarchic growth region of the disk should lower the overall eccentricity of the protoplanets there. This behavior is not immediately apparent in the top panel of figure \ref{fig:fulldisk_e_m}. There are, however, a couple of factors in the wide disk simulations that could make this extra dynamical cooling mechanism difficult to see. Firstly, the initial eccentricity distributions of the inner and outer disk are different because of the dependence of the viscous stirring and gas drag timescales on orbital period. Additionally, the protoplanet formation timescales for the inner and outer disk are vastly different, rendering a comparison between these regions at the same moment in time inappropriate.

@@

To ensure that the boundary seen near 60 days in orbital period is not
simply a transient product of the inside-out growth throughout the
disk, we examine the time evolution of $\sigma/\Sigma$ at multiple
locations. In figure \ref{fig:pl_frac_time}, the value of
$\sigma/\Sigma$ is plotted as a function of time in 10 orbital period
bins, each with a width of 10 days.
Bins interior to 60 days are indicated with a dashed line, while those exterior are indicated with a solid line. In all radial bins, the planetesimal surface density is approaching an asymptotic value. In the inner disk, this value asymptotes to zero as the planetesimal population completely depletes. In the outer disk, dynamical friction between the embryos and planetesimals eventually throttles subsequent accretion and leaves $\sim$ 10 to 20 percent of the mass surface density as planetesimals.

@@

Next, we compare the effect of varying the initial solid surface
density profile on the resulting planetesimal and embryo
distribution. The final orbital period-eccentricity state of the
particles is shown in figure \ref{fig:surfden_profiles}, with colors
and point styles indicating the different initial surface density profiles,
and with point sizes indicating the relative masses of the bodies. There is a clear trend between solid surface density at a given orbital period and the eccentricities of the remaining planetesimals. This trend is largely set by the initial conditions, as the gravitational stirring is more vigorous when bodies are packed together closely and higher eccentricities are required to produce an equilibrating aerodynamic drag force (see equations \ref{eq:vs_timescale} and \ref{eq:ts_stokes}). Despite having significantly different masses, the planetary embryos formed in each case have remarkably similar eccentricities. This is likely due to the fact that inelastic collisions play a more significant role where the solid surface density is highest, which offsets the fact that the initial bodies started off in a dynamically hotter state.

@@

In figure \ref{fig:surfden_b}, we plot the masses of the resulting protoplanets and planetesimals in all four simulations in units of $\tilde{b}$ (see equation \ref{eq:iso}). Here, we assume that each embryo has reached its isolation mass, and that differences in mass are driven by differences in the initial surface density at the embryo's current location, along with the size of its feeding zone, which is set by the dynamics relevant at that location in the disk. By plotting the derived value of $\tilde{b}$ as a function of orbital period, differences in the dynamical interactions at different locations of the disk are made more clearly visible. As mentioned previously, $\tilde{b} = 2 \sqrt{3}$ (indicated by the horizontal dashed line) carries a special significance, being the minimum feeding zone width that a body can have. In all four simulations, the feeding zone width exceeds the minimum value in the inner disk and approaches $2 \sqrt{3}$ beyond $\sim$ 60 days. The orbital periods at which this transition occurs are remarkably similar, despite the vastly different solid surface density profiles used. This similarity indicates that the boundary between accretion modes is not dependent upon the surface density profile chosen, and also supports our conclusion that planetesimal accretion is largely complete everywhere in the disk.

>>>>

We show the final state of the ``fdHi'' simulation in figure \ref{fig:fulldisk_e_m}. In the top panel,
the initial (contours) and final (points) state of the simulation is shown in the orbital period-eccentricity plane. The size of the points indicates the relative mass of the bodies. In the bottom panel, the mass of the largest bodies is shown as a function of orbital period. The solid curves indicate the isolation mass,
\begin{equation}\label{eq:iso}
	M_{iso} = \left[ \frac{\left( 2 \pi a^2 \Sigma \tilde{b} \right)^3}{3 M_{*}} \right]^{1/2},
\end{equation}
where $\tilde{b}$ is the size of the feeding zone in units
of Hill radii. For two bodies on circular, non-inclined orbits,  $\tilde{b} = 2
\sqrt{3}$ is the smallest orbital separation that produces a non-negative Jacobi energy
and permits a close encounter (see \citet{naka88}). In typical oligarchic growth simulations \citep{kokubo98}, protoplanets tend to space themselves apart by $\tilde{b} = 10$ (although it should be noted that they do not consume all of the planetesimals within this distance).
The isolation mass is plotted in blue and orange for these two values of $\tilde{b}$, respectively.

@@

In section \ref{sec:narrow}, we postulated that the increased importance of inelastic damping in the inner, non-oligarchic growth region of the disk should lower the overall eccentricity of the protoplanets there. This behavior is not immediately apparent in the top panel of figure \ref{fig:fulldisk_e_m}. There are, however, a couple of factors in the wide disk simulations that could make this extra dynamical cooling mechanism difficult to see. Firstly, the initial eccentricity distributions of the inner and outer disk are different because of the dependence of the viscous stirring and gas drag timescales on orbital period. Additionally, the protoplanet formation timescales for the inner and outer disk are vastly different, making a comparison between these regions at the same moment in time somewhat inappropriate.

@@

To ensure that the boundary seen near 60 days in orbital period is not
simply a transient product of the inside-out growth throughout the
disk, we examine the time evolution of $\sigma/\Sigma$ at multiple
orbital periods. In figure \ref{fig:pl_frac_time}, the value of
$\sigma/\Sigma$ is plotted as a function of time in 10 orbital period
bins, each with a width of 10 days.
Bins interior to 60 days are shown by a dashed line, while those exterior are shown as a solid line. In all radial bins, the planetesimal surface density is approaching an asymptotic value. In the inner disk, this value asymptotes to zero as the planetesimal population entirely depletes. In the outer disk, dynamical friction between the embryos and planetesimals eventually throttles subsequent accretion and leaves $\sim$ 10 percent or more of the mass surface density as planetesimals. It should be noted that in a typical oligarchic growth scenario, where protoplanets space themselves apart by 10 $r_{h}$ and settle onto circular orbits (giving $\tilde{b} = 2 \sqrt{3}$), roughly 30 percent of the planetesimal mass should remain out of reach of the protoplanets.

@@

Next, we investigate how the resulting planetesimal and protoplanet distribution
changes as we vary the initial solid surface density profile.
The final orbital period-eccentricity state of the
particles in the fdHi, fdHiSteep, fdHiShallow and fdLo simulations are shown in figure \ref{fig:surfden_profiles}, with point sizes indicating the relative masses of the bodies. In all cases, the inner disk is largely depleted of planetesimals, while the outer disk contains a bimodal population of planetesimals and embryos, with a clear separation in eccentricity between the two. Despite having significantly different masses, the planetary embryos formed in all simulations have remarkably similar eccentricities. This is likely due to the fact that inelastic collisions play a more significant role where the solid surface density is highest, which offsets the fact that the initial bodies started off in a dynamically hotter state.

@@

In figure \ref{fig:surfden_b}, we plot the masses of the resulting protoplanets and planetesimals in all four simulations in units of $\tilde{b}$ (see equation \ref{eq:iso}). Here, we assume that each embryo has reached its isolation mass, and that differences in mass are driven by differences in the initial surface density at the embryo's current location, along with the size of its feeding zone, which is set by the dynamics relevant at that location in the disk. By plotting the derived value of $\tilde{b}$ as a function of orbital period, differences in the dynamical interactions at different locations of the disk are made more clearly visible. As mentioned previously, $\tilde{b} = 2 \sqrt{3}$ (indicated by the horizontal dashed line) carries a special significance, being the minimum feeding zone width that a body can have. In all four simulations, the feeding zone width exceeds the minimum value in the inner disk and approaches $2 \sqrt{3}$ beyond $\sim$ 60 days. The orbital periods at which this transition occurs are quite similar between simulations, despite the vastly different solid surface density profiles used. This indicates that the boundary between accretion modes is not dependent upon the surface density profile chosen, and also supports our conclusion that planetesimal accretion is largely complete everywhere in the disk.
